\begin{frame}{Experimental Setup}
  \begin{itemize}
    \item Initial experiments were conducted with board $6 \times 6$, $5$ mines
          (density $\approx 13.9\%$) \alert{identical} for DQN and PPO. The metric
          we aim to maximize is the \texttt{win\_rate}, defined as number of episodes
          in which the agent won over the total number of test episodes. In this first phase,
          we were able to achieve a max win rate of $\approx 80.1\%$ for DQN and $\approx 80.2\%$ for
          PPO.

          Further experiments were conducted with boards $9 \times 9$ and $10$ mines.
    \item disjoint RNG seed ranges for train / validation / held-out test
    \item final evaluation: $1000$ held-out test episodes per agent
    \item to conduct in a rigorous and ordered manner an experimental performance analysis,
          we used the \href{https://wandb.ai/site/}{Weight and Biases} platform (academic plan) and its Python Library.
  \end{itemize}
\end{frame}

\begin{frame}{Hyperparameters Tuning}
  \footnotesize
  \twocol{
    \textbf{Search pipeline (exploiting W\&B Sweeps)}
    \begin{itemize}
      \setlength{\itemsep}{2pt}
      \item \texttt{bayes} search: samples guided by past trials,
            faster convergence than random search
      \item \alert{Hyperband} Early Termination: kills worst trials at each validation
            checkpoint, saving time for promising configs
      \item \textbf{Screening} $\rightarrow$ many trials, episode
            budget scaled to board size
      \item \textbf{Promotion} $\rightarrow$ best config retrained on
            3 independent seeds ($100$k ep.) for robustness
      \item Due to time and computational constraints, our search focused on \textbf{9x9@10 mines} boards only.
    \end{itemize}
  }{
    \textbf{Search space, by parameter nature}
    \begin{itemize}
      \setlength{\itemsep}{2pt}
      \item \textit{log-uniform}: \texttt{learning\_rate}
            $\in [10^{-4}, 10^{-3}]$
      \item \textit{continuous uniform}: \\
            \texttt{clip\_epsilon} (PPO) $\in [0.1, 0.3]$, \texttt{epsilon\_beta} (DQN) $\in [0.999, 0.9998]$
      \item \textit{integer uniform}: \\
            \texttt{update\_epochs} (PPO) $\in [3, 10]$
      \item \textit{discrete grid}: \texttt{discount\_factor}
            $\in \{0.8, 0.9, 0.95, 0.99\}$,
            \texttt{replay\_buffer\_size} (DQN) $\in \{50000, 100000\}$,
      \item \textit{categorical}: \texttt{epsilon\_decay\_type} (DQN)
            $\in \{\text{linear}, \text{exponential}\}$
    \end{itemize}
  }
\end{frame}

\begin{frame}{Statistical tests: Fixing DRL algorithm, \texttt{fully\_conv} VS \texttt{global\_skip\_conv}}
  Due to the nature of the environment, the metric we aimed to maximize is the win rate.
  Statistical test used: McNemar test; it's the equivalent of the
  Wilcoxon-signed rank test for binary events. The test focuses only on disagreeing episodes.

  \small{$c = $ \texttt{num\_episodes} won by $B$ \hspace{1.5em} $b$ \texttt{num\_episodes} won by $A$.

    \texttt{p\_value} = Probability of observing $c$ wins of B over $b+c$ total victories

    $ = P(X=c)$, $X \sim \text{Binom}(n=b+c, p=0.5)$. $\mathcal{H}_0: P(A:wins, B:lose) = P(A:lose, B:wins)$}

  \textbf{DQN}
  \vspace{-0.2em}
  {\scriptsize
    \centering
    \setlength{\tabcolsep}{3pt}
    \renewcommand{\arraystretch}{0.85}

    \begin{tabular}{c|cccc}
      \texttt{agent\_seed}
         &
      Win rate (\texttt{fully\_conv})
         &
      Win rate (\texttt{global\_skip})
         &
      $p$-value
         &
      Significant ($\alpha=0.05$)
      \\
      \hline
      43 & \color{goodgreen}\textbf{69\%} & 63\% & $3.32 \times 10^{-3}$ & yes \\
      44 & 67\%                           & 67\% & $0.88$                & no  \\
      45 & 68\%                           & 67\% & $0.60$                & no
    \end{tabular}

    \par}
  \vspace{0.15em}

  \textbf{PPO}
  \vspace{-0.2em}
  {\scriptsize
    \centering
    \setlength{\tabcolsep}{3pt}
    \renewcommand{\arraystretch}{0.85}

    \begin{tabular}{c|cccc}
      \texttt{agent\_seed}
         &
      Win rate (\texttt{fully\_conv})
         &
      Win rate (\texttt{global\_skip})
         &
      $p$-value
         &
      Significant ($\alpha=0.05$)
      \\
      \hline
      43 & 62\% & \color{goodgreen}\textbf{72\%} & $7.48 \times 10^{-8}$ & yes \\
      44 & 65\% & 67\%                           & $0.28$                & no  \\
      45 & 56\% & \color{goodgreen}\textbf{65\%} & $1.68 \times 10^{-6}$ & yes
    \end{tabular}

    \par}
\end{frame}
\begin{frame}{Statistical tests: Fixing architecture, DQN vs PPO}

  \textbf{\texttt{fully-conv}}
  \vspace{-0.2em}
  {\scriptsize
    \centering
    \setlength{\tabcolsep}{3pt}
    \renewcommand{\arraystretch}{0.85}

    \begin{tabular}{c|cccc}
      \texttt{agent\_seed}
         &
      Win rate (DQN)
         &
      Win rate (PPO)
         &
      $p$-value
         &
      Significant ($\alpha=0.05$)
      \\
      \hline
      43 & \color{goodgreen}\textbf{69\%} & 62\% & $8.23 \times 10^{-4}$  & yes \\
      44 & 67\%                           & 65\% & $0.21$                 & no  \\
      45 & \color{goodgreen}\textbf{68\%} & 56\% & $8.26 \times 10^{-10}$ & yes
    \end{tabular}

    \par}
  \vspace{0.15em}

  \textbf{\texttt{global-skip}}
  \vspace{-0.2em}
  {\scriptsize
    \centering
    \setlength{\tabcolsep}{3pt}
    \renewcommand{\arraystretch}{0.85}

    \begin{tabular}{c|cccc}
      \texttt{agent\_seed}
         &
      Win rate (DQN)
         &
      Win rate (PPO)
         &
      $p$-value
         &
      Significant ($\alpha=0.05$)
      \\
      \hline
      43 & 63\% & \color{goodgreen}\textbf{72\%} & $1.52 \times 10^{-6}$ & yes \\
      44 & 67\% & 67\%                           & $1$                   & no  \\
      45 & 67\% & 65\%                           & $0.36$                & no
    \end{tabular}

    \par}

  \footnotesize{
    \textit{Observations}:
    \begin{itemize}
      \item There's a strong statistical evidence about the sinergy
            between PPO and the \texttt{global\_skip} architecture.
      \item DQN algorithm performs \textit{slightly better} by using \texttt{fully\_conv}
            architecture, even though there's not a strong statistical evidence.
    \end{itemize}
  }
  \scriptsize{
    \textbf{Note:} each test run has been performed on the same held-out test episodes: that's the reason why paired statistical tests have been used.
  }

\end{frame}
\begin{frame}{Future works}
  During this project we reasoned a lot how to generalize
  to wider boards. Basing on what we've studied until now, we put our maximum
  effort to train and use at inference time the networks independently on the board size.

  Future work might include:
  \begin{itemize}
    \item \textit{curriculum learning}:
          models learn gradually, mimicking how humans learn
          (starting from the basics and moving on to complex topics). The idea
          would be graudally adapt the learning starting from little to bigger
          and more \textit{mine dense} boards.
    \item \textit{self-attention} in global section of the \texttt{global\_skip\_connection}
          architecture. A self-attention layer would allow to each cell to measure
          the impact of all the other cells of the boards, not only to the ones of its neighboroud.
          It might preserve spatial features better than a 4x4 global average pooling.
  \end{itemize}


\end{frame}
\begin{frame}{AI acknowledgments}
  \begin{itemize}
    \item The source code interacting with W\&B to store the complete run logs was produced with the help of Claude Opus 4.6 and Gemini 3.5 Flash.
    \item The \LaTeX code in charge of producing TikZ drawings for neural architecture blocks or other drawings has been produced with the help of GPT 5.6-Sol and Gemini Flash 3.5.
  \end{itemize}
\end{frame}